Index measures for oak decline severity using phenotypic descriptors

### 2020-02-07 manuscript/manuscript.Rmd
small formatting additions

# Title: Index measures for oak decline severity using phenotypic descriptors

@@

## Abstract

@@

Oak decline syndromes are wide ranging in cause and attempts have been made to subdivide the types of decline associated with *Quercus robur* and *Q. petraea* based on the duration over which a tree has been declining and the biotic agents associated with a declining tree.
Mortality in trees suffering from COD can occur over a period of decades where as AOD can occur in 5-10 years [@denman2009].
A chronically declining oak can also subsequently begin to develop the stem bleeds associated with AOD.
The visual, manual assignment of decline status and severity in the field is variable and error prone due to the contribution of many factors.
Doing this accurately requires much expertise, experience, and training [@ferretti1997].

@@

Crown radius, canopy closure and tap tests for dead stem tissue were measured for each cardinal point.
The crown radius was measured using a vertex transponder attached to the relevant side of the stem at a height of 1.3m. Canopy closure recorded as yes or no depending on whether any other crowns are touching the sampling tree.
A high resolution photo was taken of the vase to the spring of the crown and of the stem from the base to a height of 3m for each cardinal point for later reference.

@@

Although significance was obtained between the decline status groups on both the PDI and DAI scales, Figure 5 highlights the inaccuracies in the manual decline status classifications when viewed in an individual tree context.
It can be seen that a number of the trees, when compared quantitatively using these measures, could in fact be classified with a different decline state.

@@

Plots of the individual descriptors against the decline indexes are available in Supplementary Figures 4-7. 
The following sections identify how the descriptors contribute and interact to the decline index models in specific decline cases.

### Localised descriptor contributions and trends within the PDI and DAI models

LIME identified crown volume, condition transparency and radius as the top four descriptors contributing to $M_{PDI}$ in each of the healthy, moderate and severe decline cases using lime (Table 1).
These descriptors were also the same top descriptors in the overall model importance lists for the PDI (Figure 2a).
Increasing decline severity was associated with a reduction in the size and condition of the crown.
For instance, declining trees were found to have a composite crown volume below 841m^3^.
Healthy trees were found to have a crown condition greater than 71.3%, moderately declining trees between 47.6% and 71.3% and severely declining trees a condition below 23.9%.

>>>>

# Title: Index measures for oak decline severity using phenotypic descriptors {-}

@@

## Abstract {-}

@@

Oak decline syndromes are wide ranging in cause and attempts have been made to subdivide the types of decline associated with *Quercus robur* and *Q. petraea* based on the duration over which a tree has been declining and the biotic agents associated with a declining tree.
Mortality in trees suffering from COD can occur over a period of decades whereas AOD can occur in 5-10 years [@denman2009].
A chronically declining oak can also subsequently begin to develop the stem bleeds associated with AOD.
The visual, manual assignment of decline status and severity in the field is variable and error prone due to the contribution of many factors.
Doing this accurately requires much expertise, experience, and training [@ferretti1997].

@@

Crown radius, canopy closure and tap tests for dead stem tissue were measured for each cardinal point.
The crown radius was measured using a vertex transponder attached to the relevant side of the stem at a height of 1.3m. Canopy closure recorded as yes or no depending on whether any other crowns are touching the sampling tree.
A high resolution photo was taken of the base to the top of the crown and of the stem from the base to a height of 3m for each cardinal point for later reference.

@@

Although significance was obtained between the decline status groups on both the PDI and DAI scales, Figure 1 highlights the inaccuracies in the manual decline status classifications when viewed in an individual tree context.
It can be seen that a number of the trees, when compared quantitatively using these measures, could in fact be classified with a different decline state.

@@

Plots of the individual descriptors against the decline indexes are available in Supplementary Figures 4-7. 
The following sections identify how the descriptors contribute and interact within the decline index models in specific decline cases.

### Local descriptor contributions and trends within the PDI and DAI models

LIME identified crown volume, condition transparency and radius as the top four descriptors contributing to $M_{PDI}$ in each of the healthy, moderate and severe decline cases (Table 1).
These descriptors were also among the same top descriptors in the overall model importance lists for the PDI (Figure 2a).
Increasing decline severity was associated with a reduction in the size and condition of the crown.
For instance, declining trees were found to have a composite crown volume below 841m^3^.
Healthy trees were found to have a crown condition greater than 71.3%, moderately declining trees between 47.6% and 71.3% and severely declining trees a condition below 23.9%.
